### 2020-10-04 manuscript/supplementary_material_validation.Rmd
add supplementary materials

# Additional Figures and Tables

(ref:plot-sim-ex) Simulated data from the HMM with the default parameters of the simulation study. Displayed are acceleration (in deg/s$^2$, multiplied by 10000), sample-to-sample (relative) angle (in rad), and velocity (in deg/s, multiplied by 10) over time (in s). Acceleration and velocity were multiplied back to their natural scale. Note that PSOs (State 3, blue dots) do not only occur after saccades because the HMM parameters were unconstrained.

```{r plot-sim-ex, fig.cap="(ref:plot-sim-ex)"}

source(here::here("simulation/model_simulation.R"))

resp.sim.true <- list(fix = list(vel = c(3, 0.35), acc = c(3, 0.15)), 
                     sac = list(vel = c(3, 10), acc = c(3, 3), angle = c(0, 1)),
                     pso = list(vel = c(3, 1), acc = c(3, 3), angle = c(pi, 1)),
                     sp = list(vel = c(3, 1), acc = c(3, 0.15), angle = c(0, 1)))

tr.sim.true <- matrix(0.1/3, 4, 4)

diag(tr.sim.true) <- 0.9

sim.model <- simulate(HMM_simulate(n = 500, nstates = 4, trueresp = resp.sim.true, truetr = tr.sim.true,
                          truein = rep(1/4, 4)), seed = 123)

data.frame(State = as.factor(sim.model@states), Velocity = sim.model@response[[1]][[1]]@y*10,
                           Acceleration = sim.model@response[[1]][[2]]@y*1e4, Angle = sim.model@response[[1]][[3]]@y) %>% 
  mutate(Time = 1:500*0.002) %>%
  pivot_longer(c("Velocity", "Acceleration", "Angle"), names_to = "Metric", values_to = "Value") %>%
  mutate(Metric = ifelse(Metric == "Angle", "Rel. angle", Metric)) %>%
  ggplot(aes(x = Time, y = Value)) + 
  facet_grid(rows = vars(Metric), scales = "free_y", switch =  "y") + 
  geom_path() + geom_point(aes(color = State), size = 0.8) + 
  scale_y_continuous(name = "") + scale_x_continuous(name = "Time (in s)") +
  theme(strip.placement = "outside")

@@

@@

(ref:plot-schwarz-weights-task4) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 4 of the Ehinger et al. (2019) data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

@@

(ref:plot-schwarz-weights-task5) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 5 of the Ehinger et al. (2019) data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

@@

@@

(ref:plot-icl-weights-img) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the image condition of the Andersson et al. (2017) data set. Higher weights indicate better model fit.

@@

(ref:plot-icl-weights-dots) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the moving dots condition of the Andersson et al. (2017) data set. Higher weights indicate better model fit.

@@

(ref:plot-icl-weights-vid) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the video condition of the Andersson et al. (2017) data set. Higher weights indicate better model fit.

@@

(ref:plot-example-classification) Example data from Andersson et al. (2017) displayed as x-, and y-coordinates, velocity (in deg/s), acceleration (in deg/s), and sample-to-sample (relative) angle (in rad). Colors indicate classification by gazeHMM and symbols by human coder MN. Note that the human coder classified all fixations from gazeHMM as smooth pursuits and delayed the separation between the saccade and PSOs.

```{r plot-example-classification, fig.cap="(ref:plot-example-classification)", fig.height=7.5}

cbind(A2017.fit[[3]][[3]][[4]]$samples, label_MN = as.factor(A2017[[3]][[3]]$label_MN)) %>% 
  gather(c("x", "y", "vel", "acc", "angle"), key = metric, value = value, factor_key = T) %>% 
  mutate_at(vars(label, state), as.factor) %>%
  dplyr::filter(t > 4.0, t <= 4.5) %>% 
  rename(coderMN = label_MN, gazeHMM = label, Time = t) %>% 
  mutate(metric = fct_recode(metric, "x-coordinate" = "x", "y-coordinate" = "y", "Velocity" = "vel",
                             "Acceleration" = "acc", "Rel. angle" = "angle"),
         coderMN = fct_recode(coderMN, "Fix" = "1", "Sacc" = "2", "PSO" = "3", "SP" = "4", "Blink" = "5", "Other" = "6"),
         gazeHMM = fct_recode(gazeHMM, "Noise" = "0", "Fix" = "1", "Sacc" = "2", "PSO" = "3", "SP" = "4")) %>%
  ggplot(aes(x = Time, y = value)) + 
  facet_wrap(vars(metric), ncol = 1, scales = "free", strip.position = "left") +
  geom_path() + xlab("Time in s") + ylab("") +
  geom_point(aes(color = gazeHMM, shape = coderMN), size = 0.8) + 
  theme(strip.placement = "outside",
        axis.text = element_text(size = 8))

(ref:tab-example-trans) Transition Probabability Estimates for Example Data from Andersson et al. (2017)

```{r tab-example-trans}

tab.exp.tr <- cbind(as.character(1:4), matrix(round(backtrans(getpars(A2017.fit[[3]][[3]][[4]]$model))[5:20], 3), 4, 4, byrow = T))

apa_table(tab.exp.tr,
          col.names = c("From", as.character(1:4)),
          col_spanners = list("To" = c(2,5)),
          caption = "(ref:tab-example-trans)")

(ref:tab-example-resp) Response Parameter Estimates for Example Data from Andersson et al. (2017)

```{r tab-example-resp}

tab.exp.resp <- cbind(as.character(1:4), matrix(round(backtrans(getpars(A2017.fit[[3]][[3]][[4]]$model))[21:44], 2), 4, 6, byrow = T))

tab.exp.resp[1, c(6,7)] <- NA

apa_table(tab.exp.resp,
          col.names = c("State", "Shape (vel)", "Scale (vel)", "Shape (acc)", "Scale (acc)", "Mean", "Con"),
          format.args = list(na_string = "-"),
          caption = "(ref:tab-example-resp)")

>>>>

---
title             : "Keeping an Eye on Hidden Markov Models in Gaze Data Classification: Supplementary Material Validation Study"
shorttitle        : "HMMs in Gaze Classification"

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

bibliography      : "references.bib"
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r packages, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, results = "asis")

library(papaja)
library(tidyverse)
library(here)
library(psych)
library(depmixS4)
source(here("algorithm/model_helper_functions.R"))
```

```{r set graphics theme, include=FALSE}

theme_set(theme_apa())

@@

# From Wagenmakers & Farrell (2004)

schwarz.weights <- function(bic, na.rm = T) {

  d.bic <- bic - min(bic, na.rm = na.rm) # eq 2

  exp(-0.5 * d.bic)/sum(exp(-0.5 * d.bic), na.rm = na.rm) # eq 4

@@

(ref:plot-schwarz-weights-task4) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 4 of the @Ehinger2019 data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

@@

(ref:plot-schwarz-weights-task5) Schwarz weights displayed for each subject and HMMs with different numbers of states. Models were applied to task 5 of the @Ehinger2019 data set. Higher weights indicate better model fit. Grey tiles indicate erroneous model fits.

@@

load(here("validation/Andersson2017_fitted.Rdata"))

@@

(ref:plot-icl-weights-img) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the image condition of the @Andersson2017 data set. Higher weights indicate better model fit.

@@

(ref:plot-icl-weights-dots) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the moving dots condition of the @Andersson2017 data set. Higher weights indicate better model fit.

@@

(ref:plot-icl-weights-vid) ICL weights displayed for each subject and HMMs with different numbers of states. Models were applied to the video condition of the @Andersson2017 data set. Higher weights indicate better model fit.

@@

\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
